---
title: "Masters Homework 5"
author: "Katie Papciak"
date: "8/12/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 

Pull in this data, https://raw.githubusercontent.com/fredlapolla/VilcekSummerDataAnalysis2024/master/elife-32032-fig2-data1_CSV.csv, which is from this paper:

Piechura JR, Amarnath K, O'Shea EK. Natural changes in light interact with circadian regulation at promoters to control gene expression in cyanobacteria. Elife. 2017 Dec 14;6:e32032. doi: 10.7554/eLife.32032. PMID: 29239721; PMCID: PMC5785211.

Make a heatmap of the pearson's R correlation of genes that are expressed such that the rowSums are greater than 100. 

```{r}
if (!requireNamespace("ComplexHeatmap", quietly = TRUE)) {
  BiocManager::install("ComplexHeatmap")
}
library(ComplexHeatmap)
library(ggplot2)
```

```{r}
gene_exp_data<-read.csv("https://raw.githubusercontent.com/fredlapolla/VilcekSummerDataAnalysis2024/master/elife-32032-fig2-data1_CSV.csv")
gene_exp_data <- gene_exp_data[, sapply(gene_exp_data, is.numeric)]
gene_exp_filtered <- gene_exp_data[rowSums(gene_exp_data) > 100, ]
gene_exp_filtered <- gene_exp_filtered[complete.cases(gene_exp_filtered), ]
print(dim(gene_exp_filtered))
Heatmap(as.matrix(gene_exp_filtered), 
        cluster_columns = F, 
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete", 
        show_row_names = F,
        name = "Expression",
        column_title = "Samples",
        row_title = "Genes")
```



```{r}
nyc <- read.csv("https://raw.githubusercontent.com/fredlapolla/VilcekSummerDataAnalysis2024/main/NYC_HANES_DIAB.csv")
```



>- Using GGplot, make a boxplot of A1C by diabetes diagnosis. Would you assume the three groups have the same A1C level? 

```{r}
ggplot(nyc, aes(x = factor(DX_DBTS), y = A1C)) + 
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 8, size = 2, color = "white") +
  labs(title = "Boxplot of A1C by Diabetes Diagnosis",
       x = "Diabetes Diagnosis",
       y = "A1C") +
  theme_minimal()
# I would assume the pre-diabetes and diabetes diagnosed groups have higher A1C mean values given that A1C measurement is strongly positively correlated to development of diabetes.
```


>- Use ggplot2 to make a scatter plot of A1C by glucose levels.

```{r}
ggplot(nyc, aes(x = GLUCOSE, y = A1C)) +
  geom_point() +
  labs(title = "A1C vs Glucose Levels",
       x = "Glucose",
       y = "A1C") +
  theme_minimal()
```


>- Add coloration by gender for the above scatter plot and change the background theme to the minimal option. Change the X axis to label numbers by increments of 3. 

```{r}
ggplot(nyc, aes(x = GLUCOSE, y = A1C, color = factor(GENDER))) +
  geom_point() +
  labs(title = "A1C vs Glucose Levels by Gender",
       x = "Glucose",
       y = "A1C",
       color = "Gender") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(nyc$GLUCOSE, na.rm = TRUE), 
                                  max(nyc$GLUCOSE, na.rm = TRUE), 
                                  by = 3))
```

>- Create a GGPlot chart of Lead vs. Cotinine and show two different themes for the background. 

```{r}
# Theme 1: Minimal
plot1 <- ggplot(nyc, aes(x = COTININE, y = LEAD)) +
  geom_point() +
  labs(title = "Lead vs Cotinine - Minimal Theme",
       x = "Cotinine",
       y = "Lead") +
  theme_minimal()

# Theme 2: Classic
plot2 <- ggplot(nyc, aes(x = COTININE, y = LEAD)) +
  geom_point() +
  labs(title = "Lead vs Cotinine - Classic Theme",
       x = "Cotinine",
       y = "Lead") +
  theme_classic()

# Display both plots
print(plot1)
print(plot2)
```

>- Make bar charts of mean Glucose level by diabetes diagnosis. Add error bars.

```{r}
# First calculate mean and standard error for each group
library(dplyr)
glucose_summary <- nyc %>%
  group_by(DX_DBTS) %>%
  summarise(
    mean_glucose = mean(GLUCOSE, na.rm = TRUE),
    se_glucose = sd(GLUCOSE, na.rm = TRUE)/sqrt(sum(!is.na(GLUCOSE))),
    .groups = 'drop'
  )

# Create the bar chart with error bars
ggplot(glucose_summary, aes(x = factor(DX_DBTS), y = mean_glucose)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_errorbar(aes(ymin = mean_glucose - se_glucose, 
                    ymax = mean_glucose + se_glucose), 
                width = 0.2) +
  labs(title = "Mean Glucose Level by Diabetes Diagnosis",
       x = "Diabetes Diagnosis",
       y = "Mean Glucose Level") +
  theme_minimal()
```


>- Where could you go to look up other Geom shapes for different types of charts?
```{r}
# Answer: You can look up other Geom shapes in:
# 1. The ggplot2 documentation: https://ggplot2.tidyverse.org/reference/
# 2. The ggplot2 cheatsheet: https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf
# 3. The R Graphics Cookbook: https://r-graphics.org/
# 4. The ggplot2 official website: https://ggplot2.tidyverse.org
```


