---
title: "Homework1"
author: "Katie Papciak"
date: "7/9/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Homework 1

```{r, echo = TRUE}
nyc <- read.csv("https://raw.githubusercontent.com/fredlapolla/RVilcekMaster23/master/NYC_HANES_DIAB.csv")
nycdf<-data.frame(nyc)
```

>- Create a list of numbers 1 to 100 and name this list FirstHundred

```{r}
FirstHundred<-list(1:100)
print(FirstHundred)
```

>- On your own, explore nyc and find the median and mean Cadmium levels`

```{r}
print(mean(nycdf$CADMIUM, na.rm = TRUE))
print(median(nycdf$CADMIUM, na.rm = TRUE))
```


>- Find summary information about the total blood mercury levels in our NYC HANES data (hint: names(nyc) will call the names of each column). What is the max value, what is the minimum value.  

```{r}
print(summary(nycdf$MERCURYTOTALBLOOD))
```

>- Bonus/Challenge: What is the Standard Deviation and skewness of the blood mercury?

```{r}
sd_merc<- sd(nycdf$MERCURYTOTALBLOOD, na.rm = TRUE)
install.packages("moments")
library(moments)
skewness(nycdf$MERCURYTOTALBLOOD, na.rm= TRUE)
```


>- Install a package and load, such as "ggplot2" or "dplyr"

```{r}
install.packages("ggplot2")
library("ggplot2")
install.packages("dplyr")
library("dplyr")
```


>- What is a directory? How can you look up the current directory in which you are working in R?

```{r}
# A directory is the organization and hierarchy of files stored on your computer. A working directory in R is the current location where R saves and pulls files from your computer. 
getwd()
```

>- How does one read in a CSV file of data into R?

```{r}
# read.csv(file, header = TRUE, sep = ",", comment.char = "", ...) 
```

>- What two commands are needed to install and use a package in R?

```{r}
# install.packages()
# library()
```

>- What are reasons one would choose to use R for data analysis?

```{r}
# 1. Free!
# 2. Well known language that is similar to other common languages (C++, python, Java) with large user community.
# 3. Nice/customizable  data visualization and graphics packages available to public
```
